---
title: "Week 4"
author: "Lindsay Anderson"
date: "April 11, 2017"
output: html_document
---
Flourish and Blotts is looking to increase their sales by targeting specific books to display together in their store. The bookstore provided a dataset that included over 90,000 transactions of 220,000 books that are currently in their store. The goal is to determine what books to display based on preferences of customers in Oprah's book club as well as what books to display together based on books that are often purchased together. 

#Setting up the R Session and Inspecting the Data
```{r}
getwd()
setwd("C:/Users/clxa10/Desktop/Rockhurst/BIA6301/Week_4")

library(arules) 
bookbaskets <- read.transactions("bookdata.tsv.gz", format="single",  	# Note: 2 
                                 sep="\t",                    	# Note: 3 
                                 cols=c("userid", "title"),    	# Note: 4 
                                 rm.duplicates=T)       	# Note: 5

# Note 1: 
#   Load the arules package. 

# Note 2: 
#   Specify the file and the file format. 

# Note 3: 
#   Specify the column separator (a tab). 

# Note 4: 
#   Specify the column of transaction IDs and of 
#   item IDs, respectively. 

# Note 5: 
#   Tell the function to look for and remove 
#   duplicate entries (for example, multiple entries 
#   for The Hobbit by the same user). 

# 92,108 book purchases.
#220,447 user IDs.

inspect(bookbaskets[1:5]) #Examine the first five transactions
basketSizes<-size(bookbaskets) #Calculate number of books purchased by "userID"

#summary(bookbaskets)

```
The transactions above show the distribution of transactions with duplicate books. Inspecting the first five transactions shows us that the first user bought one book titled New Vegetarian: Bold and Beautiful Recipes for Every Occasion, the next three shoppers also purchased one book a piece and the fifth shopper purchased 13 books. Looking at the summary statistics tells us that the minimum number of books purchased was 1 and the maximum was 10,250 - which I would assume to be an outlier. Looking at the mean for the dataset shows us that the average number of books purchased was 11. 




##Frequently Purchased Books
The most sold books at Flouirsh and Blotts are Wild Animus at 2.7% of transactions, The Lovely Bones: A Novel at 1.4% of transactions, She's Come Undone at 1% of purchases, The Da Vinci Code for 1% of the purchases, and Harry Potter and the Sorcerer's Stone at 0.9% of the purchases. This can be seen in the barplot below of the Top 10 Purchased Book Titles.

```{r}
bookFreq<-itemFrequency(bookbaskets) #Calculate the support for each book title 
bookFreq_data_frame <- as.data.frame(itemFrequency(bookbaskets))
View(bookFreq_data_frame)
itemFrequencyPlot(bookbaskets, topN = 10, main="Top 10 Purchased Book Titles") 

bookCount <- (bookFreq/sum(bookFreq))*sum(basketSizes) #   Get the absolute count of book occurrences. 


bookbasket_use<-bookbaskets[basketSizes>1] #Only keep transactions with more than one book purchased. 

#bookbasket_use

```
#Oprah's Book Club
Flourish and Blotts would like to create a display of books similar to those available through Oprah's book club. The link to Oprah's book club is found here https://static.oprah.com/images/o2/201608/201608-obc-complete-list-01a.pdf. Most of the books that Oprah includes are novels and do not include science fiction books or mysteries. Taking a look at the top 10 purchased books in the barplot above, the books that I would look to display together based on frequency and genre would be The Lovely Bones: A Novel, She's Come Undone, The Nanny Diaries: A Novel, A Painted House, Bridget Jones's Diary, and Divine Secrets of the Ya-Ya Sisterhood: A Novel. All of these books are of the fiction and fit the profile for readers in Oprah's book club.

Let's take a look to see if there are any association rules that can be used to determine which books out those 5 suggested above should be displayed with any other books to entice buyers to purchase an additional book.


#Association Rules for Oprah's Book Club
```{r}
novelrules <- apriori(bookbaskets, parameter = list(support = 0.001, confidence = 0.1, minlen=2), appearance = list(lhs = 
                                            c("The Lovely Bones: A Novel",
                                  "She's Come Undone",
                                  "The Nanny Diaries: A Novel",
                                  "A Painted House",
                                  "Bridget Jones's Diary",
                                  "Divine Secrets of the Ya-Ya Sisterhood: A Novel"),
  default="rhs")
)
print(novelrules)
novelrules_sorted <- sort(novelrules, by="lift")
inspect(novelrules_sorted[1:10])

```
Looking at the rules suggested above, the first suggestion of displaying Bridget Jones's Diary with Bridget Jones: The Edge of Reason is already known by Flourish and Blotts because these books are in a series. The next rule shows to display Divine Secrets of the Ya-Ya Sisterhood: A novel with Little Altars Everywhere: A Novel. This is an actionable rule to follow because we have already targeted The Divine Secrets and the association rule found that if you people purchase this title they are 46 times more likely to also purchase Little Altars Everywhere.  The next pairing is The Nanny Diaries: A Novel with Confessions of a Shopaholic. This too is a good pairing as they are both fiction books that could call out to customers of Oprah's book club. People that purchase the nanny Diaries also are 34 times more likely to purchase Confessions of a Shopaholic. Finally, She's Come Undone is currently part of Oprah's book club and Vinegar Hill was part of her bookclub list in 1999. These two books being displayed together would target a present and past book for those that are nostalgic. When She's Come Undone is purchased, Vinegar Hill is 33 times more likely to be purchased as well, making this an actionable rule.  

####

# Apriori Algorithm
To answer the question regarding what books to display together, outside of those targeting Oprah's Book Club, I used the apriori method that creates association rules between all books that are purchased frequently together. I set the requirements to show only books that were purchased in at least 1% of transactions and to only show me rules that the expected results are to work 50% of the time or more. I found 200 association rules between books in this transaction set, 60 of which were 2 itemset rules, 104 were 3 itemset rules, 31 were 4 itemset rules, and 5 for 5 itemset rules. Looking through the rules, most of them are showing that books in a series are often purchased together. These series included Harry Potter, Stephanie Plum's series, Philip Pulman's series and the Lord of the Rings Trilogies. Since these are all books in a series Flourish and Blotts already displays these books together, making these rules not actionable. To find any rules that are actionable, I had to dig further down into the rules to find books that were not part of a series.
```{r}
apriori(bookbaskets) #0 Rules found
bookrules <- apriori(bookbaskets, parameter = list(support =
                                                      0.001, confidence = 0.5, minlen = 2)) 
print(bookrules) #200 rules found

#summary(bookrules)

#2-itemset: 60 rules

#3-itemset: 104 rules

#4-itemset: 31 rules

#5-itemset: 5 rules

#Total = 200 rules.

inspect(bookrules[1:10]) #Looking at the first 10 rules, not in any specific order

#Sort the rules by Lift
bookrules_sorted <- sort(bookrules, by = "lift")
#inspect(bookrules_sorted[1:10])

#Sorting by LIft and Confidence
bookrules_sorted <-sort(bookrules, by = c("lift", "confidence"))
inspect(bookrules_sorted[1:20])

#Plots of the bookrules by Support and Confidence
library(arulesViz)
plot(bookrules)

plot(bookrules, shading="order", control=list(main="Two-key plot")) #The two-key plot shows the number of association rules from rules with 2 itemsets (books) to rules containing 5 itemsets. 

#plot(bookrules, method="grouped")


plot(bookrules_sorted[1:10], method="graph") #As you can see in the graph of 20 rules, all of these associations are based on books in a series.

#plot(bookrules_sorted[1:20], method="graph", control=list(type="itemsets"))

```

The scatter plot above for 200 rules shows the support - how often the rule happens versus the confidence - how often the rule is correct. We should look for rules in the middle of the x-axis and towards the top of the y-axis. The two-key plot then takes this scatter a step further and shows which of the rules include 2 books, 3 books, 4 books or 5 books. Finally, the graph of the 10 rules shows which books are purchased together based on the support and confidence described. 

#Looking into the Association Rules
```{r}

#Removing Wild Animus from the dataset as this book appears frequently due to the author making it available for free at bookstores. This will book will not help Flourish and Blotts sell books. 
bookbaskets_noAnimus <- apriori(bookbaskets, parameter = list(support = 0.001, confidence = 0.5, minlen = 2, maxlen = 10),appearance = list(none = c("Wild Animus")))
#inspect(bookbaskets_noAnimus[1:10]) #The first 10 rules are all series so I will inspect the next set of rules to find books not in a series.
#inspect(bookbaskets_noAnimus[11:20]) #Looking at the next set of rules these are also all books in series.
#inspect(bookbaskets_noAnimus[21:70]) #Also all books in series

#Sort the rules by Lift
noAnimus_sorted <- sort(bookbaskets_noAnimus, by = "lift")
#inspect(noAnimus_sorted[1:10])
inspect(noAnimus_sorted[150:175]) 

write(bookrules_sorted, file = "C:/Users/clxa10/Desktop/Rockhurst/BIA6301/Week_4/bookrules_sorted.csv",
      sep = ",", quote = TRUE, row.names = FALSE) #write the rules to a file for sharing with Flouirsh and Blotts


```

##Recommendation
I looked through all of the rules to find any books that are bought together but are not part of a series. One set of books that are frequently bought together but are not in a series include The Pelican Brief, The Rainmaker (0.1% of purchases each) and The Client. People who buy the Pelican Brief and the Rainmaker are 122 times more likely to also purchase The client. This rule has a confidence of 63%, which could give you a significant boost in book sales. These books are all by the author John Grisham but are not part of a series. Similarly, people who purchase The Chamber and The Pelican Brief (0.1% of purchases) are 116 times more likely to purchase A Time to Kill. Displaying The Chamber and The Pelican Brief together with A Time to Kill should result in the purchase of all three, 69% of the time. I would suggest displaying all of these John Grisham books together in order to entice buyers to purchase more than one of his books at a time.
